Calculating Entropies With Nested Sampling

### 2017-01-17 paper/ms.tex
Working out differential entropy complications.

\section{Nested Sampling}
Nested Sampling \citep[NS][]{skilling2006nested}

\subsection{The sequence of $X$ values}
When $N=1$, the sequence of $X$ values is a Poisson process.

@@

\section{Example 1: A trivial bivariate gaussian}

@@

The multimodality of the posterior here raises an interesting issue. Is
the question we really want answered ``what is the value of $T$
{\bf precisely}?'', to which the mutual information relates?
Most practicing scientists would not feel particularly informed to learn
that the vast majority of possibilities had been ruled out, if the
posterior still consisted of several widely separated modes!

Perhaps, in some applications, a more appropriate question is
``what is the value of $T$ to within $\pm$ 10\%''
{\todo or something like that.}

The priors are
\begin{align}
A           &\sim \textnormal{Exponential}(1)  \\
\log_{10} T &\sim \textnormal{Uniform}(-3, 0)  \\
\phi        &\sim \textnormal{Uniform}(0, 2\pi)
\end{align}

$A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
These results were computed using a tolerance of 0.001 for every quantity.
Results:

@@

Converting these to differential entropies, we obtain
\begin{align}
H'_{\x}     &= 1.131  \pm 0.027 \\  % max depth was 30
H'_{\d}     &= 697.75 \pm 0.31  \\  % max depth was 1000
H'_{\x, \d} &= 699.84 \pm 0.31      % max depth was 1000
\end{align}
and therefore a mutual information of {\query a negative value! Oh dear!
Something is wrong.}

>>>>

\section{Nested Sampling}

Nested Sampling \citep[NS][]{skilling2006nested}

\subsection{The sequence of $X$ values}

When $N=1$, the sequence of $X$ values is a Poisson process.

@@

\subsection{Discrete vs. differential entropies}

Continuous hypothesis spaces are commonly used in practice. Suppose
$\x$ has probability density $f(\x)$. The algorithm actually estimates
\begin{align}
Q &= \int f(\x) \int f(\x')
            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
                        \, d\x' \, d\x.
\end{align}
If we actually want the differential
entropy
\begin{align}
h_{\x} &= -\int f(x) \log f(x) \, dx
\end{align}
we can use the fact that density equals mass divided
by volume. Assume $\epsilon$ is small, so that
\begin{align}
P(d(\x', \x) < \epsilon | \x)
    &\approx
    f(\x) \int_{d(\x', \x) < \epsilon} \, d\x'\\
    &= f(\x) \, V
\end{align}
where $V$ is the volume of the region where $d(\x', \x) < \epsilon$.

\section{Example 1: Measuring the mean of a gaussian}

@@

The multimodality of the posterior here raises an interesting issue. Is
the question we really want answered ``what is the value of $T$
{\bf precisely}?'', to which the mutual information relates?
Most practicing scientists would not feel particularly informed to learn
that the vast majority of possibilities had been ruled out, if the
posterior still consisted of several widely separated modes!
Perhaps, in some applications, a more appropriate question is
``what is the value of $T$ to within $\pm$ 10\%''
{\todo or something like that.}

The priors are
\begin{align}
A           &\sim \textnormal{Exponential}(1)  \\
\log_{10} T &\sim \textnormal{Uniform}(-3, 0)  \\
\phi        &\sim \textnormal{Uniform}(0, 2\pi)
\end{align}
$A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
These results were computed using a tolerance of 0.001 for every quantity.
Results:

@@

Converting these to differential entropies, we obtain
\begin{align}
H'_{\x}     &= 1.131  \pm 0.027 \\  % max depth was 30
H'_{\d}     &= 697.75 \pm 0.31  \\  % max depth was 1000
H'_{\x, \d} &= 699.84 \pm 0.31      % max depth was 1000
\end{align}
and therefore a mutual information of {\query a negative value! Oh dear!
Something is wrong. Update: well I found a bug after this, so it might
change. :)}
